%!TEX root = ../main.tex
\section{Evaluation of generic software development tools} % (fold)
\label{sec:evaluation_of_generic_software_development_tools}
In this chapter, we will look at some of the different tools used for GSD, and specify a set of tools we will focus our following investigation on.

\subsection{Currently used GSD tools} % (fold)
\label{sub:currently_used_gsd_tools}
In this section we will try to list some of the tools, currently used for GSD. Due to the large amount of different tools available, we do not aim to make an exhaustive list. Instead we will focus our list of tools associated with the popular Agile method for development , as well as a section of well used traditional tools, used for software development.

There is two reasons for focusing on Agile development tools. The first is the popularity of using this method has in GSD. The second is that using agile principles with KBE, is a practice currently being developed and utilized buy AkerSolutions KbeDesign. The benefits of such tools have already been studied by both AkerSolutions and Lars Barlindhaug in his thesis \citep{aunit}, and the fact that tools for agile development is currently being in development, makes interesting to study the possibility of integrating ready-made solutions from the world of GSD.

Agile development has introduced a lot of new development tools. Tools like unit test frameworks, which enables the developer to run the same tests over and over again. Mock frameworks, which simplifies and standardizes the test writing. Test runners, which allows tests to be automated \citep{agile_samurai} \citep{Martin:2008:CCH:1388398}. Alongside tools for testing, agile projects often utilize tools for continuous integration (CI). CI is the process of automating the building and testing process in software development, and in this thesis Lars Barlinghaug showed the feasibility and relevance of such practices in an KBE environment \citep{aunit}.

Although Agile development is currently receiving a lot of attention in the GSD community, the use of support tools in GSD is not limited to the tools associated with agile development. One of the most overlooked tool is the editor or IDE the developer uses to write the code. The general practice in GSD is that several editor options exist for every programming language, all though some might be better suited than others.

Debugging tools, which allow the developer to step through the execution of the program and inspect variable instantiation, is a common support tool. The value of being able to perform such inspections on a KBE model

The use of previously written modules and libraries, is also a practice, widely used in GSD. Using previously written code, makes development faster and less error prone. This is not a new concept for KBE development, indeed AML facilitates the creation of reusable modules and libraries to a great extent. Would it be possible to also use modules and libraries from the larger GSD community? This could hypothetically extend the choices of the KBE developer, something that makes this an interesting case to study.

IDE or editor used to write the code is an essential tool. This is an interesting case to study, since XEmacs currently is the only editor currently supported by the popular AML modeling framework. Could other editors of IDEs bring new features to AML?

An often overlooked tool, that all software developers use at some point, it the tools available for learning a particular software system. This is especially relevant for a KBE system, Rocca states that: the role of KBE developer should be easily undertaken by engineers and not restricted to ‘‘natural born hackers’’ \cite{rocca}. Good learning tools is a part of making this possible. This is of particular interest to the writers, as we will have to undergo training in AML as a part of the project.
% subsection currently_used_gsd_tools (end)

\subsection{Tools studied in this project} % (fold)
\label{sub:tools_studied_in_this_project}
Due to time constraints, we will not be able to study all the possible GSD tools relevant for KBE development. In this section we will give a brief argument for why we have chosen to study certain tools.

The tools for Agile development, is of particular interest, due to the investigation made by AkerSolution and Lars Barlindhaug, and we will focus on integration of previously written tools for testing.

Using modules and libraries from a GSD language is also something that we'll study. KBE development is a relatively smaller field than GSD, and as a result there is fewer ready-made components to choose from. Enabling the KBE developer to access a much larger amount of previously written components. This will require us to study the possibilities for integrating with AML from other programming languages.

Integrating AML with another editor is also something that we'll study. The lack of several supported code editing tools for AML is puzzling, as a large amount of code editing options are available for GSD. Integrating AML with another code editing tool, will also serve as a test of the strength of our findings in how to interface with AML.

Finally we'll look at the possibility of using a debugger in KBE development. A debugger is a feature which is lacking from AML, but is widely used in most of GSD.
% subsection tools_studied_in_this_project (end)
% section evaluation_of_generic_software_development_tools (end)
